Relocating a cluster of earthquakes using a single station

### 2012-07-17 paper2/paper2.tex
minor changes - improved typsetting of equations in Section 4 and fixed some figures so that they fit in a column
some minor changes

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\noindent\includegraphics[width = 20pc]{diags/ressummary_2Dsynth50eq.eps}
\caption{Statistical measures of error in the optimisation solutions for the 2D synthetic cases when all (blue)
and best (red) results are considered. The statistics $\Delta_{max}$, $\Delta_\mu$ and
$\Delta_\sigma$ are the maximum, mean and standard deviation of the coordinate error, respectively.
The bottom subplot shows the average minimum number of branches required to link the 2450 pairs.}
 \label{fig:optimisationresults-2Dsynth}
\end{figure}

@@

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\noindent\includegraphics[width = 20pc]{diags/ressummary_3Dsynth50eq.eps}
\caption{Statistical measures of error in the optimisation solutions for the 3D synthetic cases when all (blue)
and best (red) results are considered. The statistics $\Delta_{max}$, $\Delta_\mu$ and
$\Delta_\sigma$ are the maximum, mean and standard deviation of the coordinate error, respectively.
The absence of the blue and red lines below 60\% and 30\% indicates a breakdown in the solutions
when all or best optimisation result(s) are considered, respectively.}
\label{fig:optimisationresults-3Dsynth}
\end{figure}

@@

We do not propose a new technique for earthquake location using
travel time differences. Rather, we exploit the information created
by hypoDD with SVD to define a probability density (or posterior)
function
\begin{equation}
\label{eq-multi-var-Gauss-tt} P(\mathbf{e}_p|\Delta_{TT}) =
\frac{1}{(2\pi)^{\frac{3}{2}}\sqrt{|\Sigma|}} \exp
\left({-\frac{1}{2}\left([\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T
\Sigma^{-1} [\mathbf{e}_p-\mu_{\mathbf{e}_p}]\right)} \right),
\end{equation}
where
\begin{equation}
\mathbf{e}_p = (x_p,y_p,z_p)^T
\end{equation}
is the location of event $p$,
\begin{equation}
\mu_{\mathbf{e}_p} = (\mu_{x_p}, \mu_{y_p},\mu_{z_p})^T
\end{equation}
is the most likely location as determined using the travel time
data, and
\begin{equation}
\label{eq:Sigma-expression}
\Sigma = \left( \begin{array}{ccc} \sigma_{x_p}^2 & 0 & 0\\
0 &  \sigma_{y_p}^2 & 0 \\
0 & 0 & \sigma_{z_p}^2  \end{array} \right)
\end{equation}
is the covariance matrix. In this paper we define the mean location
$\mu_{\mathbf{e}_p}$ and covariance matrix by hypoDD optimum
solution and its uncertainties. It is important to note that hypoDD
must be used with SVD to obtain useful estimates of $\sigma_{x_p}$,
$\sigma_{y_p}$ and $\sigma_{z_p}$ (\textbf{ADD REF}).

We pose the location problem using the negative log likelihood
\begin{equation}
\label{eq-Lstar-tt-cwi} L(\mathbf{e}_1, \mathbf{e}_2, ...,
\mathbf{e}_1, \mathbf{e}_n) = - \sum_{i=1}^n
ln\left[P(\mathbf{e}_i|\Delta_{TT})\right] - \sum_{i=1}^{n-1}
\sum_{j=i+1}^n
ln\left[P(\delta_{CWIN}|\mathbf{e}_i,\mathbf{e}_j)\right],
\end{equation}
where $(\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n)$ is the joint
location,
\begin{equation}
\label{eq-ttcomponent} \sum_{i=1}^n
ln\left[P(\mathbf{e}_i|\Delta_{TT})\right]
\end{equation}
incorporates the travel time constraints and
\begin{equation}
\label{eq-codacomponent} \sum_{i=1}^{n-1} \sum_{j=i+1}^n
ln\left[P(\delta_{CWIN}|\mathbf{e}_i,\mathbf{e}_j)\right]
\end{equation}
 the coda waves.

We must differentiate $L$ to use the Polak-Ribiere conjugate
gradient technique of \citet{dr_Press87a}. The derivative of
$L(\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n)$ with respect to
$x_p$ is given by
\begin{equation}
\frac{\partial L}{\partial x_p} =  - \frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial x_p} \\
- \sum_{i=p+1}^{N} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_p,\mathbf{e}_i)\right]}{\partial
x_p} - \sum_{j=1}^{p-1} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_j,\mathbf{e}_p)\right]}{\partial
x_p}
\end{equation}
where $\sum_{i=p+1}^{N} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_p,\mathbf{e}_i)\right]}{\partial
x_p}$ and $\sum_{j=1}^{p-1} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_j,\mathbf{e}_p)\right]}{\partial
x_p}$ are defined in Appendix \ref{sec-Appendix-derivatives_ofL} and
\begin{equation}
\frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial x_p}
= -\frac{1}{2}\left([1,0,0]^T \Sigma^{-1}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}] +
[\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T \Sigma^{-1} [1,0,0] \right).
\end{equation}
Similarly, for the derivatives with respect to $y_p$ and $z_p$ we
have
\begin{equation}
\frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial y_p}
= -\frac{1}{2} \left( [0,1,0]^T \Sigma^{-1}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}] +
[\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T \Sigma^{-1} [0,1,0] \right)
\end{equation}
and
\begin{equation}
\frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial z_p}
= -\frac{1}{2} \left( [0,0,1]^T \Sigma^{-1}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}] +
[\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T \Sigma^{-1} [0,0,1] \right).
\end{equation}

@@

% This is the version using hypoDD with SVD
\begin{figure*}
\includegraphics{diags/CalaverasLoc5_hypoDD_SVD.eps}
\caption{Combined HypoDD (SVD) and CWI: all stations and 5
stations.} \label{fig-68Calaverasevents_ttandcoda1}
\end{figure*}

@@

The Omori formula
\begin{equation}
label{eq:original_omori}
 n(t) = K(t+C)^{-p}
\end{equation}
can be used to estimate the aftershock frequency per unit time
interval $n(t)$ at time $t$ (\textbf{ADD REF}). \citet{dr_Utsu95a}
converted equation \ref{eq:original_omori} to a cumulated number of
events
\begin{equation}
N(t) = K \frac{c^{1-p} + (t+c)^{1-p}}{p-1}.
\end{equation}

@@

% This is the version using hypoDD with SVD
\begin{figure*}
\includegraphics{diags/CalaverasLoc6_hypoDD_SVD.eps}
\caption{Mimicking the deployment of a temporary network by ignoring
data from all but station CCO for 50\% (or 34) of the events using
by combining CWI and TT (left) and using TT only (right). Only by
combining the data is it possible to locate all 68 events.
Furthermore, combining the data leads to a solution more consistent
with figure ..... } \label{fig-68Calaverasevents_ttsubsetandcoda1}
\end{figure*}
%%=========================================================================

>>>>

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\noindent\includegraphics[width = 20pc]{diags/synth2Dmulti/ressummary_2Dsynth50eq.eps}
\caption{Statistical measures of error in the optimisation solutions for the 2D synthetic cases when all (grey)
and best (black) results are considered. The statistics $\Delta_{max}$, $\Delta_\mu$ and
$\Delta_\sigma$ are the maximum, mean and standard deviation of the coordinate error, respectively.
The bottom subplot shows the average minimum number of branches required to link the 2450 pairs.}
 \label{fig:optimisationresults-2Dsynth}
\end{figure}

@@

\begin{figure}
%\noindent\includegraphics[width = 20pc]{../../thesis_version2/diags/eq_location_optimisation/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\noindent\includegraphics[width = 20pc]{diags/synth3Dmulti/ressummary_3Dsynth50eq.eps}
\caption{Statistical measures of error in the optimisation solutions for the 3D synthetic cases when all (grey)
and best (black) results are considered. The statistics $\Delta_{max}$, $\Delta_\mu$ and
$\Delta_\sigma$ are the maximum, mean and standard deviation of the coordinate error, respectively.
The absence of the blue and red lines below 60\% and 30\% indicates a breakdown in the solutions
when all or best optimisation result(s) are considered, respectively.}
\label{fig:optimisationresults-3Dsynth}
\end{figure}

@@

We do not propose a new technique for earthquake location using
travel time differences. Rather, we exploit the information created
by hypoDD with SVD to define a probability density (or posterior)
function
\begin{equation}
\label{eq-multi-var-Gauss-tt} 
\begin{array}{l}
P(\mathbf{e}_p|\Delta_{TT}) 
\frac{1}{(2\pi)^{\frac{3}{2}}\sqrt{|\Sigma|}} \\
\hspace{5em} \times \exp
\left({-\frac{1}{2}\left([\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T
\Sigma^{-1} [\mathbf{e}_p-\mu_{\mathbf{e}_p}]\right)} \right),
\end{array}
\end{equation}
where
\begin{equation}
\mathbf{e}_p = (x_p,y_p,z_p)^T
\end{equation}
is the location of event $p$,
\begin{equation}
\mu_{\mathbf{e}_p} = (\mu_{x_p}, \mu_{y_p},\mu_{z_p})^T
\end{equation}
is the most likely location as determined using the travel time
data, and
\begin{equation}
\label{eq:Sigma-expression}
\Sigma = \left( \begin{array}{ccc} \sigma_{x_p}^2 & 0 & 0\\
0 &  \sigma_{y_p}^2 & 0 \\
0 & 0 & \sigma_{z_p}^2  \end{array} \right)
\end{equation}
is the covariance matrix. In this paper we define the mean location
$\mu_{\mathbf{e}_p}$ and covariance matrix by hypoDD optimum
solution and its uncertainties. It is important to note that hypoDD
must be used with SVD to obtain useful estimates of $\sigma_{x_p}$,
$\sigma_{y_p}$ and $\sigma_{z_p}$ (\textbf{ADD REF}).

We pose the location problem using the negative log likelihood
\begin{equation}
\label{eq-Lstar-tt-cwi}
\begin{array}{l}
 L(\mathbf{e}_1, \mathbf{e}_2, ...,
\mathbf{e}_1, \mathbf{e}_n) = - \sum_{i=1}^n
ln\left[P(\mathbf{e}_i|\Delta_{TT})\right] \\
\hspace{6em}  - \sum_{i=1}^{n-1}
\sum_{j=i+1}^n
ln\left[P(\delta_{CWIN}|\mathbf{e}_i,\mathbf{e}_j)\right],
\end{array}
\end{equation}
where $(\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n)$ is the joint
location,
\begin{equation}
\label{eq-ttcomponent} \sum_{i=1}^n
ln\left[P(\mathbf{e}_i|\Delta_{TT})\right]
\end{equation}
incorporates the travel time constraints and
\begin{equation}
\label{eq-codacomponent} \sum_{i=1}^{n-1} \sum_{j=i+1}^n
ln\left[P(\delta_{CWIN}|\mathbf{e}_i,\mathbf{e}_j)\right]
\end{equation}
 the coda waves.

We must differentiate $L$ to use the Polak-Ribiere conjugate
gradient technique of \citet{dr_Press87a}. The derivative of
$L(\mathbf{e}_1, \mathbf{e}_2, ..., \mathbf{e}_n)$ with respect to
$x_p$ is given by
\begin{equation}
\begin{array}{l}
\frac{\partial L}{\partial x_p} =  - \frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial x_p}
- \sum_{i=p+1}^{N} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_p,\mathbf{e}_i)\right]}{\partial
x_p} \\
\hspace{5em} - \sum_{j=1}^{p-1} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_j,\mathbf{e}_p)\right]}{\partial
x_p}
\end{array}
\end{equation}
where 
\begin{equation}
\sum_{i=p+1}^{N} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_p,\mathbf{e}_i)\right]}{\partial
x_p} 
\end{equation}
and 
\begin{equation}
\sum_{j=1}^{p-1} \frac{ \partial \ln
\left[P(\delta_{CWIN}|\mathbf{e}_j,\mathbf{e}_p)\right]}{\partial
x_p} 
\end{equation}
are defined in Appendix \ref{sec-Appendix-derivatives_ofL} and
\begin{equation}
\begin{array}{l}
\frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial x_p}
= -\frac{1}{2}[1,0,0]^T \Sigma^{-1}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}] \\
\hspace{9em} -\frac{1}{2}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T \Sigma^{-1} [1,0,0].
\end{array}
\end{equation}
Similarly, for the derivatives with respect to $y_p$ and $z_p$ we
have
\begin{equation}
\begin{array}{l}
\frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial y_p}
= -\frac{1}{2} [0,1,0]^T \Sigma^{-1}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}] \\
\hspace{9em} -\frac{1}{2}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T \Sigma^{-1} [0,1,0] 
\end{array}
\end{equation}
and
\begin{equation}
\begin{array}{l}
\frac{\partial ln\left[P(\mathbf{e}_p|t_{DD})\right]}{\partial z_p}
= -\frac{1}{2} [0,0,1]^T \Sigma^{-1}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}] \\
\hspace{9em} -\frac{1}{2}
[\mathbf{e}_p-\mu_{\mathbf{e}_p}]^T \Sigma^{-1} [0,0,1].
\end{array}
\end{equation}

@@

% This is the version using hypoDD with SVD
\begin{figure}
\includegraphics{diags/CalaverasLoc5_hypoDD_SVD.eps}
\caption{Combined HypoDD (SVD) and CWI: all stations and 5
stations.} \label{fig-68Calaverasevents_ttandcoda1}
\end{figure}

@@

The Omori formula
\begin{equation}
\label{eq:original_omori}
 n(t) = K(t+C)^{-p}
\end{equation}
can be used to estimate the aftershock frequency per unit time
interval $n(t)$ at time $t$ (\textbf{ADD REF}). \citet{dr_Utsu95a}
converted equation \ref{eq:original_omori} to a cumulated number of
events
\begin{equation}
N(t) = K \frac{c^{1-p} + (t+c)^{1-p}}{p-1}.
\end{equation}

@@

% This is the version using hypoDD with SVD
\begin{figure}
\includegraphics{diags/CalaverasLoc6_hypoDD_SVD.eps}
\caption{Mimicking the deployment of a temporary network by ignoring
data from all but station CCO for 50\% (or 34) of the events using
by combining CWI and TT (left) and using TT only (right). Only by
combining the data is it possible to locate all 68 events.
Furthermore, combining the data leads to a solution more consistent
with figure ..... } \label{fig-68Calaverasevents_ttsubsetandcoda1}
\end{figure}
%%=========================================================================
